SMT-Boosted Security Types for Low-Level MPC

### 2025-01-30 esop25/camera-ready/examples.tex
checkpoint

In Section \ref{section-smt} we gave the example of 3-party additive
sharing for any field $\fieldp{p}$. In Sections \ref{section-ipj} and
\ref{section-metalangty} we discussed the example of YGC input encoding.
Now we consider examples that illustrate the versatility of the
$\TirName{Encode}$ typing rule for confidentiality, the use of
postconditions for correctness, and the use of both pre-~and
postconditions for integrity.

To evaluate our system especially regarding performance, using the
Satisfiability Modulo Fields theory in cvc5 on an Apple M1 processor,
we have verified correctness of additive sharing protocols in fields
up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit integers)
which takes $\sim$7ms to compute for 3 parties and not significantly more
than for, e.g., $\fieldp{2}$. We have also verified the semi-homomorphic
encryption properties leveraged in the BDOZ circuit library functions
(Figures \ref{fig-bdozsum} and \ref{fig-bdozmult}) in fields up to
size $\fieldp{2^{31} - 1}$ which take $\sim$8ms for the property used for
the postcondition of the $\ttt{mult}$ gate. This is categorically
better performance than can be achieved using previous brute-force
methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.

@@

\paragraph{GMW Input Encoding and And Gate}
In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
binary circuit protocol, values are encrypted in a manner similar to
BDOZ as described in Section \ref{section-ipj}. In our definition of
GMW we use the convention that shared values $\macgv{\mesg{w}}$ are
identified by strings $w$ and encoded as shares $\mx{w}{1}$ and
$\mx{w}{2}$.  As for YGC, ciphertypes reflect the confidentiality of
GMW input encodings as defined in the $\ttt{encodegmw}$ function in
Figure \ref{fig-gmw}. No programmer annotation is needed given the
syntax of the function body. More interestingly, the $\ttt{andgmw}$
gate definition shows how a programmer hint can express the relevant
confidentiality property of the output share $\mx{z}{2}$ on client 2,
using the $\ttt{as}$ casting. The non-trivial equivalence can be
verified by SMT once during verification, and subsequently
confidentiality is expressed in the dependent type of the function (as
for $\ttt{encode}$).

>>>>

In Section \ref{section-smt} we gave the example of 3-party additive
sharing for any field $\fieldp{p}$. In Sections \ref{section-ipj} and
\ref{section-metalangty} we discussed the example of YGC input encoding.
Now we consider examples that illustrate the versatility of the
$\TirName{Encode}$ typing rule for confidentiality and the use of
postconditions for correctness. In an extended version of this paper
we also exemplify the use of both pre-~and
postconditions for integrity \cite{skalka2025smtboostedsecuritytypeslowlevel}.

To evaluate our system especially regarding performance, using the
Satisfiability Modulo Fields theory in cvc5 on an Apple M1 processor,
we have verified correctness of additive sharing protocols in fields
up to size $\fieldp{2^{31} - 1}$ (an approximation of 32-bit integers)
which takes $\sim$7ms to compute for 3 parties and not significantly more
than for, e.g., $\fieldp{2}$. We have also verified the semi-homomorphic
encryption properties leveraged in the BDOZ circuit library functions
defined in \cite{skalka2025smtboostedsecuritytypeslowlevel} in fields up to
size $\fieldp{2^{31} - 1}$ which take $\sim$8ms for the property used for
the postcondition of the $\ttt{mult}$ gate. This is categorically
better performance than can be achieved using previous brute-force
methods in $\fieldp{2}$ \cite{skalka-near-ppdp24}.

@@

\paragraph{GMW Input Encoding and And Gate.}
In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
binary circuit protocol, values are encrypted in a manner similar to
using secret sharing as described in Sections
\ref{section-smt-additive-sharing} and \ref{section-ipj}. In our
definition of GMW we use the convention that shared values
$\macgv{\mesg{w}}$ are identified by strings $w$ and encoded as shares
$\mx{w}{1}$ and $\mx{w}{2}$.  Ciphertypes reflect the
confidentiality of GMW input encodings as defined in the
$\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No programmer
annotation is needed given the syntax of the function body. More
interestingly, the $\ttt{andgmw}$ gate definition shows how a
programmer hint can express the relevant confidentiality property of
the output share $\mx{z}{2}$ on client 2, using the $\ttt{as}$
casting. The non-trivial equivalence can be verified by SMT once
during verification, and subsequently confidentiality is expressed in
the dependent type of the function (as for $\ttt{encode}$).

### 2025-01-30 esop25/camera-ready/ipj.tex
checkpoint

In the above example, note that the environment implicitly records
that $\rvl{w}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. An implementation
of a malicious-secure arithmetic circuit that
applies this rule is provided in extended version of
this paper \ref{skalka2025smtboostedsecuritytypeslowlevel}.

>>>>

In the above example, note that the environment implicitly records
that $\rvl{w}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. We implement
a malicious-secure arithmetic circuit that
applies this rule in extended version of
this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.

### 2025-01-30 esop25/camera-ready/metalang.tex
checkpoint

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which are guaranteed to
have closed types, to obtain Hoare typings for instructions. In Figure
\ref{fig-mtj} we define the type derivation rules for $\metaprot$. In
the \TirName{Mesg} rule we use algorithmic confidentiality and
integrity type checking in a straightforward manner.  In the
\TirName{Encode} rule, we verify the hint given in the type
annotation, and then use it for type checking.  In the \TirName{App}
rule we verify preconditions of the given function type and
instantiate the type and postconditions with the given expression
arguments.  The \TirName{Sig} rule we specify how to verify $\Pi$
types for functions. Note that the values $\mv_1,\ldots,\mv_n$ chosen
to instantiate the function parameters are chosen ``fresh'', meaning
that any strings and client identifiers occurring as substrings in
$\mv_1,\ldots,\mv_n$ are unique and not used in any program source
code. This ensures generality of verification as demonstrated in Lemma
\ref{lemma-eqs-notg} in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.

>>>>

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which are guaranteed to
have closed types, to obtain Hoare typings for instructions. In Figure
\ref{fig-mtj} we define the type derivation rules for $\metaprot$. In
the \TirName{Mesg} rule we use algorithmic confidentiality and
integrity type checking in a straightforward manner.  In the
\TirName{Encode} rule, we verify the hint given in the type
annotation, and then use it for type checking.  In the \TirName{App}
rule we verify preconditions of the given function type and
instantiate the type and postconditions with the given expression
arguments.  The \TirName{Sig} rule we specify how to verify $\Pi$
types for functions. Note that the values $\mv_1,\ldots,\mv_n$ chosen
to instantiate the function parameters are chosen ``fresh'', meaning
that any strings and client identifiers occurring as substrings in
$\mv_1,\ldots,\mv_n$ are unique and not used in any program source
code. This ensures generality of verification as demonstrated
in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.

### 2025-01-30 esop25/camera-ready/smt.tex
checkpoint

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party additive secret sharing protocol, where
each party generates three additive, information theoretically secure
secret shares by subtracting random field values used as
one-time-pads. Two shares are communicated to the other clients and
one remains local. This technique is information-theoretically secure
\cite{evans2018pragmatic} due to the one-time use of random
values-- that is, $< 3$ shares reveal nothing about the
secret, but the sum of all three \emph{is} the secret. Each party
publicly reveals the sums all their shares and then outputs the sum of the reveals.
{\footnotesize
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$}
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

>>>>

\subsubsection{Example: Correctness of 3-Party Addition}
\label{section-smt-additive-sharing}
Consider the following 3-party additive secret sharing protocol, where
each party generates three additive, information theoretically secure
secret shares by subtracting random field values used as
one-time-pads. Two shares are communicated to the other clients and
one remains local. This technique is information-theoretically secure
\cite{evans2018pragmatic} due to the one-time use of random
values-- that is, $< 3$ shares reveal nothing about the
secret, but the sum of all three \emph{is} the secret. Each party
publicly reveals the sums all their shares and then outputs the sum of the reveals.
{\footnotesize
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$}
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$
